
\section{Results and Conclusion}
\label{sec:results}

All hypotheses are mainly confirmed by the empirical findings:

\begin{itemize}
  \item Through their code contribution firm employed developers have a strong positive impact on the code contribution of external developers
  \item Firm employed developers have a strong positive impact by participating on issues and comments in issues by causing more participation of external users
  \item The first year seems to be the most important time period for firm employed developers' participation: High contribution ratio in the beginning benefits popularity and success of the OS project in the long run
  \item The higher the overall share of contribution from firm employed developers, the more likely the project is successful and accepted by the OS community
\end{itemize}

Nevertheless, influence of higher participation of firm employed developers could not be verified ultimately on some continuous social metrics \footnote{\textit{Forks} and \textit{Number of Issues} on some models} which might be caused by broad difference of popularity and level of awareness between firms and software segments.

Beside the regression results we also receive interesting fringe insights: According to the frequency of commits with respect to daytime we can assume, that most firms and OS developers do work in USA, Canada and Europe. Also do external developers seem to be employed as well. Most developers commit source code at the usual working time period (9 a.m. - 5 p.m.) and "external developers" do not differ significantly from as "firm employed" classified developers. This would imply, that there is a stronger collaboration between firms than between firms and "libertine" developers.

\subsection{Implications and Managerial Advice}

Firms should initiate and invest \footnote{with funds and man-power} in their OS projects. The share of code contribution by firm employees should be very high - especially in the beginning - to achieve a broad acceptance by the OS community and developers on (rival) firms \footnote{contribution in the beginning can be interpreted as groundwork that needs to be done by firm employed developers to build a solid software foundation}. The return of investment will be a multiplier effect on commitment \footnote{via source code contribution and participation in communication channels} by external developers and a gain on level of awareness. Beside that, firms' OS commitment might lead to more cross-firm collaboration, improve employment of external developers and bring forward establishing new technologies in smaller amounts of time. If the share of firm employed developers is higher than usual, projects seem more likely to reach a certain degree of quality, maturity and attention. As result the community is willing to contribute their professional expertise for free to the project.

% As the selection of firms (by their number of popular projects) reveals, almost all well-known and favored software technology firms are represented.

\subsection{Further Research}
\label{sec:further_research}

The empirical data implies further research with respect to broader observations of firms and projects. By analyzing more firms and projects the hypotheses could eventually be applied more generally to software technology driven firms. Consideration of financial data of firms could reveal the impact of OS commitment on firms' revenue. And more sophisticated employees firm rating would help to investigate the coherence of OS commitment and quality of work environment. Furthermore, observations of network relations between firms, developers and users would improve the understanding of the importance of collaboration and exchange of knowledge.
%
% \begin{itemize}
%   \item observe all firms
%   \item firm financial data
%   \item firm / work place ratings by employees (developers)
%   \item collaboration between firm developers
%   \item measure firm reputation in open source communities (improves more OS commitment firm's reputation in the community / for potential employees)
% \end{itemize}

% \subsection{Outlook and Discussion}
%
% \begin{itemize}
%   \item further research (network effects, cross firm collaboration)
% \end{itemize}

% TODO: Results, Conclusion, Discussion


% Help tp Lontables
% Source: http://tex.stackexchange.com/questions/11380/how-to-repeat-top-rows-column-headings-on-every-page
% \documentclass{article}
% \usepackage{longtable}
% \begin{document}
% \begin{longtable}{ccc}
% \hline
% Column 1 & Column 2 & Column 3\\
% \hline
% \endhead % all the lines above this will be repeated on every page
% A & B & C\\
% A & B & C\\
% ... many lines of table ...
% \end{longtable}
% \end{document}
% --> \endhead
